# Time Series Tools

In this module, we will be learning the tools that will allow us to analyze Time Series data. Before any forecasts are estimated, it is essential to get data ready for analysis. Additionally, we should understand the data structure and any general patterns the data might be exhibiting. This will inform us of the type of analysis we should perform on the data.  

We will first learn how to get the data ready for analysis using `dplyr` and `tidyr`. Secondly, we will plot the data to identify the trend, seasonality, or cyclical behavior. To graph our time series, we will be learning the `ggplot2` package in R. Lastly, since time series deals with dates, we will introduce a package called `lubridate`, that helps us parse and manipulate dates.

## The Avocado Data Set

The avocado data is weekly retail scan data for U.S retail volume (units) and price. Retail scan data comes directly from retailers' cash registers based on actual retail sales of Hass avocados. The data reflects an expanded, multi-outlet retail data set. Multi-outlet reporting includes an aggregation of the following channels: grocery, mass, club, drug, dollar, and military. The Average Price (of avocados) in the table reflects a per unit (per avocado) cost, even when multiple units (avocados) are sold in bags. The Product Lookup codes (PLUs) in the data are only for Hass avocados. Other avocados (e.g. greenskins) are not included in this data.

## Data Wrangling With tidyverse.

`tidyverse` is a collection of packages in R that allow us to manipulate, explore and visualize data. There are a couple of packages within tidyverse (`dplyr` and `tidyr`) that we will be using to transform our data and get it ready for analysis. `dplyr` will allow us to do most of our data manipulation: creating new variables, renaming variables, filtering values, sorting, grouping, and summarizing, among others. `tidyr` will allow us to pivot data sets, unite or separate columns, and deal with missing values. Although it is always possible to complete these tasks using base R, `tidyverse` allows us to efficiently perform these operations using data manipulation verbs that are very intuitive to the user. Below we load the library.

```{r}
library(tidyverse)
```

As you can see, several packages were attached (loaded) when we write  `library(tidyverse)`. As mentioned, both `tidyr` and `dplyr` are part of this overall package. Now that the package is loaded we can import our data by using the `read_csv()` function from the `readr` package.

```{r}
avocado<-read_csv("https://jagelves.github.io/Data/avocado2020.csv")
```

The function imports the data as a tibble (a data structure similar to a data frame). There are three variables that are classified as character, while the rest are double. At this point you can preview the data with either the `spec()` or `glimpse()` commands.

You will notice that the *date* variable is of type character. You can convince yourself of this by using the `class()` function:

```{r}
class(avocado$date)
```

To coerce this variable to a date we can use the `lubridate` package. Specifically, since the *date* variable is formatted as month/day/year we will use the `mdy()` function.

```{r}
library(lubridate, warn.conflicts = F, quietly = T)
avocado$date<-mdy(avocado$date)
```

We can confirm that the type of the variable has now been coerced:

```{r}
class(avocado$date)
```

When using `dplyr` it's always helpful to use piping. Generally speaking, piping allows us to chain functions. Piping (`%>%`) passes the object on the left of the pipe as the first argument to the right of the pipe. We can illustrate this by using the `select()` and `arrange()` functions.

```{r}
avocado %>% select(c(average_price,geography)) %>%
  arrange(desc(average_price)) 
```

```{r}
avocado %>% filter(date!=ymd("2018-01-01")) -> avocado
``` 
There is a lot to unpack in this line of code. Let's start with the functions used. Both the `select()` and `arrange()` functions are part of the `dplyr` package. As the name indicates, the `select()` function selects variables from a tibble or data frame. The `arrange()` function sorts the data. By default it will sort in ascending order, hence we have used the `desc()` function to use descending order.

Now, let's focus on the entire code by reading it from left to right. *avocado* is the tibble that contains all of the data. Since it is to the left of the pipe (%>%), it passes as the first argument of the `select()` function. That is why you don't see *avocado* as the first argument listed in the `select()` function. The new data frame (i.e., the one with only the geography and the average price) then passes as the first argument of the `arrange()` function that follows the second pipe. That data frame is sorted in descending order so that the highest average avocado price is displayed first.

This example highlights the use of `dplyr` functions to transform your data. There are plenty of other functions you can use, but learning these are outside the scope of this book. To find out more we recommend reading @R4DS chapter 4. For now we will use one more data transformation technique to retrieve average price of organic avocados for California.


```{r}
avocado %>% 
  filter(geography=="California", type=="organic",
         year<=2018) %>%
  select(date, average_price, geography) -> cali
```

Whereas the `select()` function chooses particular variables, the `filter()` function chooses rows of the tibble that meet the conditions listed.

## Visualizing The Data

To visualize the data we will be using `ggplot2`. One of the main functions in `ggplot2` is the `aes()` function. This function sets the plotting canvas and determines which variables are to be plotted. The `geom_line()` function specifies the type of plot. In time series we will use the line plot regularly. Below you can see the code to create a line plot of the average price of avocados for California.

```{r}
ggplot(data=cali) + 
  geom_line(mapping=aes(x=date,y=average_price, group=geography),color="black") +
  theme_classic() + labs(x="",y="Average Price", 
                         title="Organic Avocado Price in California",
                         subtitle="2015-2018")  
```

The graph shows that the average price of avocados in California has been increasing. The average price reached a maximum of about 2.6 in 2017 and was at a minimum in in the spring of 2015. There seems to me also some seasonal patterns with low prices at the beginning of the year and peaks at the end.

## tssibbles and Decomposition

```{r}
library(fpp3, quietly = T, warn.conflicts = F)
avocado %>%
  as_tsibble(key=c(type, geography),
           index=date, regular=T) %>%
  filter_index("2015-01-04"~"2018-12-31")-> avocadots
```

```{r}
avocadots %>% filter(geography=="California", type=="organic") %>%
  select(date,geography,average_price,total_volume) -> calits

calits  %>% mutate(
  `10-MA`=slider::slide_dbl(average_price,mean,
                            .before=5, .after=4,.complete=T)) -> calits

calits %>%
  autoplot(average_price)+
  geom_line(aes(y=`10-MA`), col="red")
```

```{r}
calits  %>% mutate(
  `2X10-MA`=slider::slide_dbl(`10-MA`,mean,
                           .before=1, .after=0,.complete=T)) -> calits

calits %>%
  autoplot(average_price)+
  geom_line(aes(y=`2X10-MA`), col="red")

```

```{r}
calits %>% filter_index("2015-01-04"~"2018-12-31") %>%
  model(STL(average_price~trend(window=150)+
              season(window=52), robust=TRUE)) %>%
  components() %>% autoplot()+ theme_classic()

```
## Chipotle Wants You to Forecast Avocado Prices

Chipotle is an American chain specializing in tacos and burritos that are made to order in front of the customer. Guacamole is the perfect pairing to their delicious food and one of Chipotle's best sellers. Their guac uses just six ingredients: avocados, lime juice, cilantro, red onion, jalapeño, and kosher salt. Because of its popularity, each restaurant goes through approximately five cases of avocados a day, amounting to more than 44,000 pounds of avocados annually. Chipotle wants you to develop a model to forecast the price of avocados. This model will allow the company to

## Lessons Learned in This Chapter

In this module you have been introduced to data wrangling etc.

-   renaming, selecting and filtering variables.

-   plotting using ggplot.

-   handling dates with lubridate.

## Readings

@FPP3 chapter 1 and chapter 2.\
tidy data: https://vita.had.co.nz/papers/tidy-data.pdf
